---
title: "Lab 6"
author: "Marco Palombo"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE, messages = FALSE, warnings = FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warnings = FALSE)

# attach packages
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (or medians)
H1: ranks are not equal

### Creat samples 

`set.seed()` to create "pseudorandom, but reproducible samples"
```{r}
set.seed(1414)
gp_1 <-  sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

```{r}
hist(gp_1)
hist(gp_2)
```

try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1,gp_2)

#if these data were paired, I can just add the argument 'paired = TRUE' within the t.test() fucntion
```

This p-value means that IF our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value,2)` probability of finding sample means *at least as different* as mine by random chance.

[A more comprehensive explanation of difference between groups...actually means, data spread, effect size]...there is no significant difference (t(df) = t-statistic value, p = ).

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that IF our samples are drawn from populations with the same median (i.e. if the null hypothesis is true), there is `r round(mwu$p.value,2)` probability of finding sample medians *at least as different* as mine by random chance.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic` p = `r round(mwu$p.value,2)`)

## Simple linear regression

Two *continuous variables* that you are trying to understand the relationship between and you think that a linear relationship makes sense.

Use the `pengins` data from {pamerpenguiuns}, exploring flipper length (x) versus body  mass (y)

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = 'lm')
  

```

### find the model estimates (coefficients)

Use the `lm` function for linear model.

```{r}
penguins_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

Model coefficients: `r penguins_lm$coefficients`
Flipper length coefficient: `r penguins_lm$coefficients[2]`

### Use `broomLLtidy()` to get a tidy output from lm()

```{r}
penguins_lm_tidy <- tidy(penguins_lm)

penguins_lm_tidy
```

### Use broom::glance() to get output of model fit and statistical significance

```{r}
penguins_lm_glance <- glance(penguins_lm)

penguins_lm_glance
```

### use equatiomatic::extract_eq() to return the model in LaTeX

```{r}
extract_eq(model = penguins_lm, use_coefs = TRUE)
```

